---
title: "Speed Dating Analysis"
author: "Tamara Makarova"
date: "8 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Speed Dating Exploratory Data Analysis

```{r}
library(ggplot2)
library(reshape)
```


```{r echo=FALSE}
setwd("Documents/Analysis/Speed Dating Analysis")
data <- read.csv("data/Speed Dating Data.csv")
dim(data)
```

```{r}
ggplot(data, aes(x=race_o)) + geom_bar()
ggplot(data, aes(x=age_o)) + geom_histogram()
summary(data$age_o)
by(data$age_o, data$gender,  summary)

head(data[data$iid==1])
data[data$iid==1, c("iid", "id", "idg", "gender", "order", "partner", "pid", "match", "position", "positin1")]
summary(data[data$iid <=20, c("idg")])
```

Primary goal in participating in speed dating

```{r}
#Get goal data for each participant 
agg_fun <- function(arr) {
  return (arr[1])
}

survey <- aggregate(data[,c("iid", "gender", "age", "field", "race", "goal", 
                           "date", "go_out", "career")], by=list(data$iid), 
                    FUN = agg_fun)

survey$goal <- as.factor(survey$goal)
survey$gender <- as.factor(survey$gender)
ggplot(survey, aes(x=goal, y=..count../sum(count))) +
  geom_bar(na.rm = FALSE) +
  scale_x_discrete(breaks=c("1","2","3","4","5","6","NA"),
                   labels = c("Fun night out", "Meet new people", "get a date",
                              "serious rel", "Say I did it", "Other", "NA")) +
  coord_flip()

#????How to calculate percentages in each group, what is NA? 
ggplot(survey, aes(x=goal)) +
  geom_bar(aes(y=..count../sum(..count..), fill=gender), position = position_dodge()) +
  scale_x_discrete(breaks=c("1","2","3","4","5","6","NA"),
                   labels = c("Fun night out", "Meet new people", "get a date",
                              "serious rel", "Say I did it", "Other", "NA")) +
  scale_fill_hue(labels=c("women", "men")) +
  scale_y_continuous(labels = scales::percent ) +
  coord_flip()
```

How often do people go on a date? Go out?

```{r}
survey$date <- as.factor(survey$date)
survey$go_out <- as.factor(survey$go_out)

ggplot(survey, aes(x=date, y=..count../sum(count))) +
  geom_bar() +
  scale_x_discrete(breaks=c("1","2","3","4","5","6","7"))

ggplot(survey, aes(x=go_out, y=..count../sum(count))) +
  geom_bar() +
  scale_x_discrete(breaks=c("1","2","3","4","5","6","7"))

```

What do woman think about men preferrencies? Is it true?

```{r}
# aggregate data

library(ggplot2)
library(reshape)

setwd("Documents/Analysis/Speed Dating Analysis")
data <- read.csv("data/Speed Dating Data.csv")

keep <- c("attr1_1", "sinc1_1", "intel1_1", 
                               "fun1_1", "amb1_1", "shar1_1",
                               "attr2_1", "sinc2_1", "intel2_1", 
                               "fun2_1", "amb2_1", "shar2_1")

agg_fun <- function(arr) { return (arr[1]) }
pref <- melt (aggregate(data[,keep], by=data[c("iid", "gender", "age")], 
                       FUN = agg_fun),
              id = c("iid", "gender", "age"))

get_attrib <- function (attrib) {
  atstr <- as.character(attrib)
  patt <- "[0-9]_[0-9]$"
  pos <- regexpr(patt, atstr)[1]
  if (pos >= 1)
    return (substr(atstr, 1, pos-1))
  else 
    return (NA)
}  

get_attrib_type <- function (attrib) {
  atstr <- as.character(attrib)
  patt <- "[0-9]_[0-9]$"
  pos <- regexpr(patt, atstr)[1]
  if (pos >= 1)
    return (substr(atstr, pos, nchar(atstr)))
  else
    return (NA)
  
}  

pref$attrib <- get_attrib(pref$variable)
pref$attrib_type <- get_attrib_type(pref$variable)

```



```{r}
head(pref)
levels(as.factor(pref$attrib))

levels(as.factor(pref$attrib_type))


gender_pref <- aggregate(pref[c("value")], 
                         by = pref[c("gender", "attrib", "attrib_type")],
                         function(x) median(x, na.rm = TRUE))

gender_pref

ggplot(pref, aes(x=attr1_1)) + geom_histogram()
ggplot(pref, aes(x=sinc1_1)) + geom_histogram()
ggplot(pref, aes(x=intel1_1)) + geom_histogram()
ggplot(pref, aes(x=fun1_1)) + geom_histogram()

pref[,c("attr1_1", "sinc1_1", "intel1_1", 
                               "fun1_1", "amb1_1", "shar1_1")]

```





```{r}

data$match <- as.factor(data$match)
ggplot(data, aes(x=age, y=age_o) ) + geom_point(alpha=0.5) + facet_wrap(~match)
```

Personal estimate vs real estimate.
We need to calcualte average estimate from other participants

```{r}
outside.est <- aggregate(data[,c("iid", "gender", "age",
                                "attr3_1", "sinc3_1", "intel3_1", 
                               "fun3_1", "amb3_1", 
                               "attr_o", "sinc_o", "intel_o", 
                               "fun_o", "amb_o")],
                    by = list(data$iid),
                    function (x) mean(x, na.rm = TRUE))

ggplot(outside.est, aes(x=attr3_1, y=attr_o)) + 
  geom_point(position="jitter", alpha=0.5) + 
  geom_abline() +
  scale_y_continuous(limits=c(0,10)) +
  ggtitle("Attractivness")

ggplot(outside.est, aes(x=sinc3_1, y=sinc_o)) + 
  geom_point(position="jitter", alpha=0.5) + 
  geom_abline() +
  scale_y_continuous(limits=c(2.5,10)) +
  ggtitle("Sincere")

ggplot(outside.est, aes(x=intel3_1, y=intel_o)) + 
  geom_point(position="jitter", alpha=0.5) + 
  geom_abline() +
  scale_y_continuous(limits=c(2.5,10)) +
  ggtitle("Intelligent")

ggplot(outside.est, aes(x=fun3_1, y=fun_o)) + 
  geom_point(position="jitter", alpha=0.5) + 
  geom_abline() +
  scale_y_continuous(limits=c(0,10)) +
  ggtitle("Fun")

ggplot(outside.est, aes(x=amb3_1, y=amb_o)) + 
  geom_point(position="jitter", alpha=0.5) + 
  geom_abline() +
  scale_y_continuous(limits=c(0,10)) +
  ggtitle("Ambitious")


```

In average people seem to overestimate themselves.
Only for attract and fun variation in estimates is quite high. For all other characteristics people tend 
to give average estimates (In order to see these characteristics in a person much more time  is needed)


Do we know what we want?

```{r}
keep <- c("attr_o", "sinc_o", "intel_o", 
                               "fun_o", "amb_o", "shar_o")

liked <- melt (aggregate(data[, keep], by=list("pid"), 
                        subset = data$dec_o==1,
                        FUN = agg_fun),
              id = c("pid"))


```

